\documentclass[12pt,a4paper]{article}

% ─── Packages 
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

% ─── Colors 
\definecolor{primaryblue}{RGB}{30, 80, 160}
\definecolor{lightgray}{RGB}{245, 245, 245}
\definecolor{darkgray}{RGB}{80, 80, 80}

% ─── Hyperref Setup 
\hypersetup{
    colorlinks=true,
    linkcolor=primaryblue,
    urlcolor=primaryblue,
    pdfauthor={RISCO Team},
    pdftitle={RISCO – System Documentation}
}

% ─── Header / Footer 
\pagestyle{fancy}
\fancyhf{}
\rhead{\textcolor{darkgray}{\small RISCO – System Documentation}}
\lhead{\textcolor{darkgray}{\small Task 1}}
\rfoot{\textcolor{darkgray}{\small Page \thepage}}
\renewcommand{\headrulewidth}{0.4pt}

% ─── Section Formatting 
\titleformat{\section}
  {\Large\bfseries\color{primaryblue}}
  {\thesection.}{0.5em}{}[\titlerule]

\titleformat{\subsection}
  {\large\bfseries\color{darkgray}}
  {\thesubsection.}{0.5em}{}

% ─── Document 
\begin{document}

% ── Title Page 
\begin{titlepage}
    \centering
    \vspace*{2cm}

    {\Huge\bfseries\color{primaryblue} RISCO}\\[0.4cm]
    {\Large\color{darkgray} Financial Risk Management Platform}\\[0.2cm]
    \rule{\linewidth}{0.5pt}\\[1cm]

    {\LARGE\bfseries System Documentation}\\[0.3cm]
    {\large Task 1 – SCD Project Report}\\[2cm]

    \begin{tabular}{ll}
        \textbf{Course}   & Software Construction and Development \\[4pt]
        \textbf{Document} & System Overview \& Requirements \\[4pt]
        \textbf{Version}  & 1.0 \\
    \end{tabular}

    \vspace{1.5cm}

    {\large\bfseries\color{darkgray} Group Members}\\[0.4cm]
    \begin{tabular}{ll}
        Inshal Ejaz   & SE-24053 \\[4pt]
        Nabiha Rizwan & SE-24054 \\[4pt]
        Erma Khan     & SE-24060 \\[4pt]
        Ali Raza      & SE-24302 \\
    \end{tabular}

    \vfill
    {\small\color{darkgray} Prepared as part of the SCD CEA submission.}
\end{titlepage}

% ── Table of Contents 
\tableofcontents
\newpage


\section{System Overview}


\textbf{RISCO} (Risk Intelligence \& Strategic Cash Operations) is a relational
database-driven financial risk management platform designed specifically for
startups. The system addresses a critical gap in existing financial tools:
the tendency to report on historical data \emph{after} problems have occurred,
rather than identifying and preventing risks in advance.

\subsection{Problem Statement}

Startups frequently fail due to unseen financial risks—delayed client payments,
sudden expense spikes, and gradual cost creep. Traditional finance applications
focus on historical reporting and detect issues only after they have already
affected the organization's financial health.

\subsection{Proposed Solution}

RISCO shifts financial management from \emph{reactive reporting} to
\emph{proactive risk prevention}. The platform stores organizational financial
data in structured monthly snapshots—capturing revenue, expenses, cash balance,
and liabilities—and uses this data to:

\begin{itemize}[leftmargin=1.5cm]
    \item Calculate key indicators such as \textbf{burn rate} and
          \textbf{cash runway}.
    \item Analyze trends over time to detect cost creep, expense spikes, and
          revenue inconsistencies.
    \item Forecast the future cash-flow impact of current financial patterns.
    \item Generate risk alerts ranked by urgency to guide decision-making.
\end{itemize}

The system is built on a normalized relational database model that organizes
users, organizations, roles, financial snapshots, clients, revenue streams, and
expense records. This structure ensures secure multi-user access while enabling
complex analytical queries for forecasting and risk modelling.

\section{System Users}

RISCO supports \textbf{role-based multi-user access}. Each user belongs to one
or more organizations with a specific role that controls what they can view or
modify. The four predefined roles are described in Table~\ref{tab:roles}.

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{>{\bfseries}p{3cm} c p{7.5cm}}
\toprule
Role & Permission Level & Description \\
\midrule
Owner       & 3 – Full Access  & Creates the organization; has unrestricted
                                  access to all features, including user
                                  management and data deletion. \\
Co-Owner    & 3 – Full Access  & Shares full administrative rights with the
                                  Owner; can manage members and all financial
                                  records. \\
Accountant  & 2 – Edit / Add   & Can add and update financial snapshots,
                                  revenue entries, and expense records, but
                                  cannot manage organization membership. \\
Stakeholder & 1 – View Only    & Read-only access to dashboards, reports, and
                                  risk alerts; cannot modify any data. \\
\bottomrule
\end{tabular}
\caption{System roles and permission levels}
\label{tab:roles}
\end{table}

A single \textbf{User} account can belong to multiple organizations
simultaneously, each with a different role assignment.

\section{Key Features of the System}

\subsection{Financial Snapshot Monitoring}
RISCO records a monthly financial snapshot per organization containing the cash
balance, total liabilities, monthly revenue, and monthly expenses. A unique
constraint on \texttt{(snapshot\_date, organization\_id)} prevents duplicate
entries and ensures data integrity.

\subsection{Burn Rate \& Cash Runway Calculation}
From snapshot data, the system automatically calculates:
\begin{itemize}[leftmargin=1.5cm]
    \item \textbf{Burn Rate} – the net monthly cash outflow when expenses
          exceed revenue.
    \item \textbf{Cash Runway} – the estimated number of months until the
          organization runs out of funds at the current burn rate.
\end{itemize}

\subsection{Revenue Intelligence}
The platform tracks both \emph{recurring} and \emph{one-time} revenue entries
linked to individual clients. It monitors expected versus actual payment dates
to detect late payments and assign a \textbf{client reliability score} that
reflects historical payment behavior.

\subsection{Expense Intelligence}
Expenses are categorized by urgency (\emph{Critical} or \emph{Non-Critical})
and type (\emph{Recurring} or \emph{One-Time}). This categorization enables
the system to prioritize cost-reduction recommendations and flag abnormal
expense spikes.

\subsection{Automated Risk Alerts}
The system generates \texttt{RiskAlerts} linked to specific revenue or expense
records. Each alert carries:
\begin{itemize}[leftmargin=1.5cm]
    \item An urgency level: \textbf{Low (1)}, \textbf{Medium (2)}, or
          \textbf{Critical (3)}.
    \item A status lifecycle: \emph{Open} $\to$ \emph{In Progress} $\to$
          \emph{Resolved} / \emph{Ignored}.
    \item A resolution deadline to encourage timely action.
\end{itemize}

\subsection{Forecasting Module}
A dedicated \texttt{Forecasting} table links snapshots to forward-looking
projections, allowing the system to estimate future financial risk based on
current trends and historical patterns.

\subsection{Role-Based Access Control (RBAC)}
All features are gated by the user's assigned role within an organization,
ensuring that sensitive financial data is accessible only to authorized
personnel.


\section{Functional Requirements}


The following functional requirements define the core operations the system
must support.

\subsection{User \& Authentication Management}
\begin{enumerate}[label=FR-\arabic*., leftmargin=2cm]
    \item The system shall allow users to register with a unique email address
          and a securely hashed password.
    \item The system shall authenticate users via email and password at login.
    \item The system shall support password hashing to protect user credentials
          at rest.
\end{enumerate}

\subsection{Organization \& Membership Management}
\begin{enumerate}[label=FR-\arabic*., leftmargin=2cm, start=4]
    \item The system shall allow authenticated users to create a new
          organization, automatically assigning them the \emph{Owner} role.
    \item The system shall allow Owners and Co-Owners to add other users to
          their organization with a specified role.
    \item The system shall enforce role-based permissions so that users can
          only perform actions permitted by their assigned role.
    \item The system shall support one user belonging to multiple organizations
          simultaneously.
\end{enumerate}

\subsection{Financial Data Entry}
\begin{enumerate}[label=FR-\arabic*., leftmargin=2cm, start=8]
    \item The system shall allow authorized users to create and update monthly
          financial snapshots containing cash balance, liabilities, revenue,
          and expenses.
    \item The system shall enforce a unique constraint to prevent more than one
          snapshot per organization per month.
    \item The system shall allow authorized users to add revenue entries
          specifying the client, revenue type, expected date, received date,
          and amount.
    \item The system shall allow authorized users to add expense entries
          specifying urgency, expense type, date, and amount.
\end{enumerate}

\subsection{Client Management}
\begin{enumerate}[label=FR-\arabic*., leftmargin=2cm, start=12]
    \item The system shall allow authorized users to register clients with
          contact information linked to their organization.
    \item The system shall calculate and store a \textbf{reliability score}
          for each client based on historical payment punctuality.
\end{enumerate}

\subsection{Risk \& Analytics}
\begin{enumerate}[label=FR-\arabic*., leftmargin=2cm, start=14]
    \item The system shall calculate the \textbf{burn rate} from monthly
          snapshot data.
    \item The system shall calculate the \textbf{cash runway} based on the
          current cash balance and burn rate.
    \item The system shall automatically generate risk alerts when predefined
          financial thresholds are breached.
    \item The system shall allow users to update the status of a risk alert
          through its lifecycle (\emph{Open} $\to$ \emph{Resolved} /
          \emph{Ignored}).
    \item The system shall produce forecasting records linking financial
          snapshots to future projections.
\end{enumerate}


\section{Summary}

RISCO is a proactive, database-driven financial risk management platform
purpose-built for startups. Unlike conventional finance applications such as
QuickBooks or Xero---which focus on historical reporting---RISCO is designed
to \emph{predict} financial risk before problems escalate into crises. The
platform addresses the most common causes of early-stage startup failure:
delayed client payments, sudden expense spikes, and gradual cost creep. By
storing monthly financial snapshots, the system derives critical indicators
including \textbf{burn rate} and \textbf{cash runway}, while tracking
client-level revenue reliability scores and categorizing expenses by urgency
and type to support targeted cost-reduction analysis.

Risk is surfaced through an automated \texttt{RiskAlerts} system linked
directly to the revenue or expense records that triggered it, with urgency
levels and a resolution lifecycle enforcing timely intervention. A
\texttt{Forecasting} module extends this with best-case, base-case, and
worst-case projections. Access is governed by a \textbf{role-based access
control} model across four roles---Owner, Co-Owner, Accountant, and
Stakeholder---with the entire schema normalized to 4NF, ensuring data
integrity, consistency, and scalability across multiple organizations.

\end{document}
